Sensitivity of Simulations of Double Detonation Type Ia Supernova to Integration Methodology

### 2023-09-02 subch_tols/paper.tex
a bit more res
overleaf version

\documentclass[modern]{aastex631}

@@

\newcommand{\maestro}{{\sf MAESTRO}}
\newcommand{\castro}{{\sf Castro}}
\newcommand{\amrex}{{\sf AMReX}}
\newcommand{\pynucastro}{{\sf pynucastro}}

@@

@@

\begin{document}
%======================================================================
% Title
%======================================================================
\title{Influence of Time Integration Methodology on Double Detonation Type
Ia Supernova Nucleosynthesis}

\shorttitle{Integration Methods and SNe Ia Nucleosynthesis}

\author[0000-0001-8401-030X]{Michael Zingale}
\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
             Stony Brook, NY 11794-3800, USA}

@@

\author[0000-0003-0439-4556]{Max Katz}
\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
             Stony Brook, NY 11794-3800, USA}

@@

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
We study the coupling of hydrodynamics and reactions in simulations of
the double detonation model for Type Ia supernovae.  When assessing the
convergence of simulations, the focus is usually on spatial resolution, 
however, the method of coupling the physics together as well as the
tolerances used in integrating a reaction network also play an important role.
In this paper, we explore how the choices made in both coupling and integrating the reaction
portion of a simulation influences the accuracy, efficiency, and the nucleosynthesis
of simulations of the double detonation model of Type Ia supernovae.
We find that the stronger coupling between hydrodynamics and reactions,
afforded by the simplified spectral deferred corrections method introduced
previousl,y provides a more accurate and efficient result than standard
operator splitting.  We also find no need to limit reactions rates or reduce the
simulation timestep to the reaction timescale.  The entire simulation
methodology used here is freely available as part of the \castro\ simulation
code.
\end{abstract}

@@

There are many different progenitor models for Type Ia supernova (SN Ia) that
are currently being explored to explain the observed diversity of SN Ia explosions \cite{taubenberger:2017}.  The double detonation model
\citep{fink:2007} involves a sub-Chandrasekhar mass white dwarf with a
thin accreted He layer.  A detonation in the He layer propagates
around the star while sending a compression wave converging toward the
center.  This compression wave ignites a carbon detonation that then
propagates outward from the center of the white dwarf.  Many groups
have modeled this scenario, varying the initial conditions,
resolution \citep{rivas:2022}, and reaction networks to understand the robustness of this
model as well as its connection to observations.

\citep{townsley:2019} \citep{gronow:2020}, \citep{gronow:2021}, \citep{glasner:2018}, \citep{roy:2022}, \citep{shenbildsten:2014}, \cite{polin:2019}, \citep{kushnirkatz:2020}

These events are modeled using the Euler equations with reactive and
gravitational sources:
\begin{eqnarray}
\ddt{\rho} + \nabla \cdot (\rho \Ub) &=& 0 \\
\ddt{(\rho \Ub)} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &=& \rho \gb \\
\ddt{(\rho E)} + \nabla \cdot (\rho \Ub E + \Ub p) &=& \rho \Ub \cdot \gb + \rho \dot{S} \\
\ddt{(\rho X_k)} + \nabla \cdot (\rho \Ub X_k) &=& \rho \omegadot_k
\end{eqnarray}
where $\rho$ is the mass density, $\Ub$ is the velocity, $E$ is the specific
total energy, $p$ is the pressure, and $X_k$ are the mass fractions
of the nuclei that react.  The sources are the gravitational acceleration,
$\gb$, which is obtained by solving the the Poisson equation
\begin{equation}
\nabla^2 \Phi = 4 \pi G \rho
\end{equation}
(where $G$ is Newton's constant) for the gravitational potential, $\Phi$, and defining $\gb = -\nabla \Phi$.
The reaction network provides the species creation rate, $\omegadot$ and the energy generation rate, $\dot{S}$.
Finally, the system is closed via the equation of state:
\begin{equation}
p = p(\rho, e, X_k)
\end{equation}
where $e$ is the specific internal energy, obtained as $e = E - |\Ub|^2/2$.

Simulation codes use different methods to advance this system of equations.  For hydrodynamics,
explicit-in-time integration is usually used, with the timestep, $\Delta t$, restricted by the Courant limit.  In two-dimensions, this is:
\begin{equation}
\Delta t = C \min_{i,j} \left \{ \frac{\Delta x}{|u_{i,j}| + c_{i,j}}, \frac{\Delta y}{|v_{i,j}| + c_{i,j}}  \right \}
\end{equation}
where $\Delta x$ and $\Delta y$ are the grid width in the $x$- and $y$-directions, $u$ and $v$
are the $x$- and $y$-components of the velocity and $c$ is the sound
speed, and the minimum is taken over all zones on the grid.  Here, the
Courant number, $C$ is required to be $C < 1$ for stability.  We note that
this is the form of the stability criterion for the corner transport upwind
formulation \citep{ppmunsplit} of hydrodynamics we use.  Dimensionally-split
or method-of-lines formulations will have a slighly more restrictive criteria.
Reactions, on the other hand, tend to be stiff, and are usually evolved
using implicit-in-time integration, with many small steps taken to
make up the hydrodynamics timestep $\Delta t$.

@@

When using operator-splitting, the density remains constant, since there
is no reaction source in the mass continuity equation.  The mass
fractions, $X_k$, and energy evolve according to:
\begin{eqnarray}
\frac{dX_k}{dt} &= \omegadot_k(\rho, T, X_k) \label{eq:stranga} \\
\frac{de}{dt} &= \dot{S}(\rho, T, X_k) \\
T &= T(\rho, e, X_k) \label{eq:strangc}
\end{eqnarray}
In the simplest approximation of operator splitting, the
evolution of the mass fractions is solved alone, without integrating
the temperature or energy.  This is the form that was used, e.g., in FLASH
\citep{flash}.  A more accurate operator splitting also includes the
energy evolution (perhaps in terms of temperature), solving the system Eqs.~\ref{eq:stranga}--\ref{eq:strangc}.  This will require
calling the equation of state each time we evaluate the righthand side
of the ODE system, potentially increasing the computational expense.  However, as we showed in
\citet{strang_rnaas}, this is needed to get second-order convergence.

Detonations can be difficult to model numerically.  One common
approach is to use a simultion timestep based on the reaction
timescale, as discussed in \citet{prometheus}, instead of the
hydrodynamics timescale.  For the double detonation scenario, the
ignition of the second detonation can be very sensitive to how the
simulation is run \MarginPar{refs}.  Artifically limiting the
rates \MarginPar{ref}

In \citet{castro_simple_sdc}, we introduced a method based on the
ideas of spectral deferred corrections for coupling reactions and
hydrodyamics (that we called ``simplified-SDC'').  Here, the overall
time-integration is done iteratively, with the hydrodynamics seeing an
explicit reactive source, and the reaction update evolving an ODE
system that includes a piecewise-constant-in-time advective source,
$\Adv{\Uc}^{n+1/2}$,
\begin{equation}
\label{eq:simplesdc}
\ddt{\Uc} = \Rb(\Uc) + \Adv{\Uc}^{n+1/2}
\end{equation}
Here, $\Uc$ is the conserved state, $\Uc = (\rho, \rho \Ub, \rho E,
\rho X_k)^\intercal$, $\Rb(\Uc)$ are the reactive sources, and
$\Adv{\Uc}^{n+1/2}$ is an approximation to the advective update over
the timestep, as computed by the corner transport upwind formulation
of the piecewise parabolic method
\citep{ppmunsplit,millercolella:2002,ppm}. Crucially, this advective
term includes an explicit reaction source term and the update,
Eq.~\ref{eq:simplesdc} means that the reactions will ``see'' what
advection is doing as the reaction network is integrated.

In addition to the time-integration strategy, integrating the reaction
ODE system requires specifying tolerances, usually both an absolute, $\epsilon_\mathrm{abs}$, and relative tolerance, $\epsilon_\mathrm{rel}$, that are combined together into a weight of the form:
\begin{equation}
w_i = \epsilon_{\mathrm{abs},i} + \epsilon_{\mathrm{rel},i} |y_i|
\end{equation}
where the $y_i$ is one of the variables being integrated by the ODE
integrator and $i$ is the index, indicating that different tolerances
can be used for each variable.  Traditionally one set of tolerances is
used for the mass fractions and another for the energy (if it is
integrated).  Unfortunately, the tolerances used by simulations are
not normally reported in papers, and in many cases, may not be easily
controlled by a code user at runtime (and therefore, users may not be
aware of what they are using).  Some examples can be found in publicly
available networks.  For example, the freely available version of the
{\tt aprox13} network uses $\epsilon_\mathrm{rel} = 10^{-5}$ and a
floor on the mass fractions corresponding to $\epsilon_\mathrm{abs} =
10^{-6}$.  Since the simplified-SDC update works in terms of the
conserved variables, $\Uc$, we apply a density weighting to the absolute tolerances when comparing to the values used in Strang integration.

We considered an extreme version of the double detonation scenario in
\citet{castro_simple_sdc}---a very large perturbation was applied to
drive a detonation directly into the underlying carbon-oxygen white
dwarf.  We showed that the Strang split integration method had
difficulty with the integration unless we used tighter tolerances,
while the simplified-SDC method was well-behaved.  Here we look at a
more realistic setup for a double detonation, with a small
perturbation that drives a He detonation around the star and a
compression wave that results in the ignition of a C detonation near
the center of the star.  We focus on the nucleosynthesis and the
efficiency of the method.

%======================================================================
% Results
%======================================================================
\section{Simulations and Results}\label{Sec:results}

\subsection{Simulation setup}

We use the freely-available \castro\ simulation code for all the
simulations shown here.  Our domain uses a 2D axisymmetric geomety
with a size of $5.12\times 10^9~\mathrm{cm}$ by $1.024\times
10^{10}~\mathrm{cm}$.  This is much larger than the initial size of
the star, giving it plenty of room to expand (we note that the figures
shown in the results section are all zoomed in on the star).  The
coarse grid is $640\times 1280$ zones and we use 2 levels of
refinement, each a jump of $2\times$, giving a maximum resolution of
$20~\mathrm{km}$.  \castro\ uses the \amrex\ library
\citep{amrex_joss} to manage the grid hierarchy.  The refinement
strategy is picked to refine on the star and any regions where the
temperature is greater than $10^8~\mathrm{K}$.  Finally, \castro\ uses
subcycling, so finer grids are advanced at a smaller timestep than the
coarse grids.

As noted above, we use the unsplit PPM algorithm for advection, with
the Riemann solver from \citet{colellaglaz:1985} and the general
stellar equation of state of \citet{timmes_swesty:2000}.  In
all simulations, the CFL constraint is the
only timestep constraint used.  Self-gravity is done using a full
Poisson solve using geometric multigrid, with Dirichlet boundary
conditions on the domain boundary computed via a multipole expansion
with a maximum order of 6.  For the low density regions outside of the
star, we use a sponge term on the momentum equation to prevent the
very low density material that is not in hydrostatic equilibrium from
raining down on the star.

All simulations use the same reaction network: the {\tt subch\_simple}
as described in \cite{zhi2023}.  This includes 22 nuclei and 94 rates
from ReacLib \citep{reaclib} and is produced with
\pynucastro\ \citep{pynucastro2}.  \MarginPar{show I show the
  network?} Importantly, we include
$\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha,p)\isotm{Ne}{21}$
which creates protons to allow for
$\isotm{C}{12}(p,\gamma)\isotm{N}{13}(\alpha,p)\isotm{O}{16}$.  This
sequence can be faster than
$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$, as pointed out by
\citet{shenbildsten} and is important for getting the detonation speed
correctly.  To reduce the size of the network, we have
\pynucastro\ combine some $(\alpha,p)(p,\gamma)$ and $(\alpha,\gamma)$
rates into an effective $(\alpha,\gamma)$ rate.  Screening is provided
following the procedure in \citet{wallace:1982}, combining the
screening functions of \citet{graboske:1973,alastuey:1978,itoh:1979}.

The reaction system is integrated using a version of the VODE ODE
integrator \citep{vode} ported to C++.  Our modifications to this
integrator are described in \citet{castro_simple_sdc}.  We use an
analytic approximation to the Jacobian (the main approximation is that
the species derivatives of the screening function are not included)
and found that we get best results when we disable the caching of the
Jacobian in the integrator.

If during the advance of a timestep an
error is generated (negative density, ODE integration fails, mass
fractions don't sum to 1, CFL constraint violated at the new time),
the step is thrown-out and retied with a smaller timestep.  This is
done on a level-by-level basis in the overall AMR subcycling
hierarchy.

We will run a suite of simulations all using the same initial model,
constructed following the methodology in \citet{subchandra}.  We use
an isothermal core of C/O of $1.1~M_\odot$ with $T = 10^7~\mathrm{K}$
and a thin transition region where the temperature ramps up to
$1.75\times 10^8~\mathrm{K}$ and the composition changes to 99\%
\isot{He}{4} and 1\% \isot{N}{14}.  This He envelope is then
integrated isentropically.  The density of the transition from the
underlying CO white dwarf to the He envelope was selected to yield an
envelope mass of $0.05~M_\odot$.  The code for generating the model is
freely available at
\url{https://github.com/amrex-astro/initial_models}.  \MarginPar{should
  we do a zenodo release of this?}

We place a small temperature perturbation in the He layer using the
same form as in \citet{castro_simple_sdc}:
\begin{equation}
  T = T_0 \left \{ 1 + X(\isotm{He}{4}) f \left [1 + \tanh(2 - \sigma) \right ] \right \}
\end{equation}
where
\begin{equation}
  \sigma = \left [ x^2 + (y - R_0)^2 \right ]^{1/2} / \lambda
\end{equation}
and
\begin{equation}
  R_0 = R_\mathrm{pert} + R_\mathrm{base}
\end{equation}
where, $R_\mathrm{base}$ is the radius at which the helium layer
begins and $R_\mathrm{pert}$ is the distance above the base to put the
perturbation.  We choose $R_\mathrm{pert} = 100~\mathrm{km}$.  The
temperature is perturbed above the initial model value, denoted as
$T_0$ here.  The amplitude of the perturbation is $f = 3$ and the
scale of the perturbation is $\lambda = 12.5~\mathrm{km}$.  This is a
very small perturbation, but enough to seed the initial He detonation
in the envelope.  \MarginPar{give the peak temperature}

Our base simulation uses the simplified-SDC integration, a CFL number
of $0.2$, and sets $\epsilon_\mathrm{rel} = \epsilon_\mathrm{abs} =
10^{-5}$ for both the species and energy.  Overall, the simulation
requires 21,784 coarse grid timesteps to evolve to $1.0~\mathrm{s}$ of
simulation time.

The simulations are all run on the OLCF Frontier machine, using 4
nodes / 32 AMD GPUs.  The data is moved to the GPUs at the start of
the simulation and all computation is done there.  Our GPU offloading
strategy \citep{castro_gpu} takes advantage of the \amrex\ C++
lambda-capturing functionality to be performance portable.
All of the simulation code is on github at
\url{https://github.com/amrex-astro/} and the inputs files, global
diagnostics used to make the line plots shown below, and metadata
describing the git hashes, compiler environment, and runtime
parameters for all of the simulaitons is available at \MarginPar{add
  Zenodo ref}.

>>>>

\documentclass[preprint,linenumbers]{aastex631}

@@

\newcommand{\maestro}{{\sf MAESTRO}}
\newcommand{\castro}{{\sf Castro}}
\newcommand{\amrex}{{\sf AMReX}}
\newcommand{\pynucastro}{{\sf pynucastro}}
\newcommand{\microphysics}{{\sf Microphysics}}
\newcommand{\flash}{{\sf Flash}}

@@

\newcommand{\add}[1]{{\color{red} #1}}

@@

\begin{document}
%======================================================================
% Title
%======================================================================
\title{Sensitivity of Simulations of Double Detonation Type
Ia Supernova to Integration Methodology}

\shorttitle{Double Detonation SNe Ia and Integration Methods}

\author[0000-0001-8401-030X]{Michael Zingale}
\affiliation{Department of Physics and Astronomy, Stony Brook University,
             Stony Brook, NY 11794-3800, USA}

@@

\author[0000-0002-0297-0313]{Melissa Rasmussen}
\affiliation{Department of Physics and Astronomy,
Stony Brook University,
Stony Brook, NY 11794-3800, USA}

\author[0000-0002-1633-6495]{Abigail Polin}
\affiliation{The Observatories of the Carnegie Institution for Science, 813 Santa Barbara St., Pasadena, CA 91101, USA}
\affiliation{TAPIR, Walter Burke Institute for Theoretical Physics, 350-17, Caltech, Pasadena, CA 91125, USA}

\author[0000-0003-0439-4556]{Max Katz}
\affiliation{Department of Physics and Astronomy, Stony Brook University,
             Stony Brook, NY 11794-3800, USA}

@@

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
We study the coupling of hydrodynamics and reactions in simulations of the double detonation model for Type Ia supernovae.  When assessing the
convergence of simulations, the focus is usually on spatial resolution; 
however, the method of coupling the physics together as well as the
tolerances used in integrating a reaction network also play an important role.
In this paper, we explore how the choices made in both coupling and integrating the reaction
portion of a simulation (operator / Strang splitting vs.\ the simplified spectral deferred corrections method we introduced previously) influences the accuracy, efficiency, and the nucleosynthesis
of simulations of double detonations.
We find no need to limit reaction rates or reduce the
simulation timestep to the reaction timescale.  The entire simulation
methodology used here is GPU-accelerated and made freely available as part of the \castro\ simulation
code.
\end{abstract}

@@

Many different progenitor models for Type Ia supernovae (SNe Ia) 
are currently being explored to explain the observed spectroscopy diversity of SNe Ia explosions \citep{taubenberger:2017}. One model, the double detonation model
\citep{NOMOTO82,WOO_WEAV84}, involves a sub-Chandrasekhar mass white dwarf with an accreted He layer. A detonation in the $\mathrm{He}$ layer propagates
around the star while sending a compression wave converging toward the
center.  This compression wave ignites a carbon detonation that then
propagates outward from the center of the white dwarf, releasing enough
nuclear binding energy to gravitationally unbind the star.

Many simulations have been performed on this model, varying the initial conditions, reaction network sizes, and resolution. Recent works show
that, particularly in the case of a thin helium shell, a double detonation is viable progenitor for some SNe Ia \citep{fink:2007,mollwoosley:2013,shenbildsten:2014,glasner:2018,shen:2018,polin:2019,townsley:2019,kushnirkatz:2020,gronow:2020,gronow:2021,boos:2021,roy:2022,rivas:2022}.  These works use slightly different methods of integrating the reaction network and controlling the timestep, which we discuss in the sections below.  Our goal is to understand how sensitive a simulation of double detonations is to the details of how the reaction
network is integrated and how it is coupled to hydrodynamics.

All of these works solve the compressible Euler equations with reactive and
gravitational sources:
\begin{eqnarray}
\label{eq:euler1}
\ddt{\rho} + \nabla \cdot (\rho \Ub) &=& 0 \\
\ddt{(\rho \Ub)} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &=& \rho \gb \\
\ddt{(\rho E)} + \nabla \cdot (\rho \Ub E + \Ub p) &=& \rho \Ub \cdot \gb + \rho \dot{S} \\
\label{eq:euler4}
\ddt{(\rho X_k)} + \nabla \cdot (\rho \Ub X_k) &=& \rho \omegadot_k
\end{eqnarray}
where $\rho$ is the mass density, $\Ub$ is the velocity, $E$ is the specific
total energy, $p$ is the pressure, and $X_k$ are the mass fractions
of the nuclei that react.  The gravitational acceleration,
$\gb$, is obtained by solving the Poisson equation,
\begin{equation}
\nabla^2 \Phi = 4 \pi G \rho \enskip ,
\end{equation}
(where $G$ is Newton's constant) for the gravitational potential, $\Phi$, and defining $\gb = -\nabla \Phi$.  We note that $\Phi$ is often approximated as a monopole or via a multipole expansion, which works well so long as the star
stays reasonably spherically symmetric.
The reaction network provides the species creation rate, $\omegadot_k$, and the energy generation rate, $\dot{S}$.
Finally, the system is closed via the equation of state:
\begin{eqnarray}
p &= p(\rho, e, X_k) \\
T &= T(\rho, e, X_k)
\end{eqnarray}
where $e$ is the specific internal energy, obtained as $e = E - |\Ub|^2/2$.

This is a multiphysics system of equations and the different physical processes have different associated timescales.
Correspondingly, simulation codes use different methods to advance this system of equations.  For hydrodynamics,
explicit-in-time integration is usually used, with the timestep, $\Delta t$, restricted by the Courant limit (the precise form depends on whether we are doing
unsplit hydrodynamics,  dimensionally-split, or method of lines time-integration, see \citealt{ppmunsplit}).
%In two-dimensions, this is: \MarginPar{I thinks we can skip this! -alex}
%\begin{equation}
%\Delta t = C \min_{i,j} \left \{ \frac{\Delta x}{|u_{i,j}| + c_{i,j}}, \frac{\Delta y}{|v_{i,j}| + c_{i,j}}  \right \}
%\end{equation}
%where $\Delta x$ and $\Delta y$ are the grid width in the $x$- and $y$-directions, $u$ and $v$
%are the $x$- and $y$-components of the velocity and $c$ is the sound
%speed, and the minimum is taken over all zones on the grid.  Here, the
%Courant number, $C$ is required to be $C < 1$ for stability.  We note that
%this is the form of the stability criterion for the corner transport upwind
%formulation \citep{ppmunsplit} of hydrodynamics we use here.  Dimensionally-split
%or method-of-lines formulations will have a different, slightly more restrictive %criterion.
In contrast, reactions are usually evolved
using implicit-in-time integration, with many small steps taken to
make up the hydrodynamics timestep $\Delta t$.

@@

When using operator splitting, the density remains constant during the reactive update, since there
is no reaction source in the mass continuity equation.  The mass
fractions, $X_k$, and energy evolve according to:
\begin{eqnarray}
\frac{dX_k}{dt} &= \omegadot_k(\rho, T, X_k) \label{eq:stranga} \\
\frac{de}{dt} &= \dot{S}(\rho, T, X_k)  \label{eq:strang_energy}\\
T &= T(\rho, e, X_k) \label{eq:strangc}
\end{eqnarray}
In the simplest approximation of operator splitting, the
evolution of the mass fractions is solved alone, without integrating
the temperature or energy, Eq.~\ref{eq:strang_energy}.  This is the form that is used, e.g., in \flash\
(\citealt{flash}, see also \citealt{townsley:2016}).  A more accurate operator splitting also includes the
energy evolution (perhaps in terms of temperature), solving the system Eqs.~\ref{eq:stranga}--\ref{eq:strangc}.  This is the default in \castro~\citep{castro,castro_joss},
and temperature evolution is also used in the works of \citet{gronow:2020,gronow:2021} following the implementation of \citet{pakmor:2012}. This evolution requires
calling the equation of state each time we evaluate the right-hand side
of the ODE system, potentially increasing the computational expense.  However, as we showed in
\citet{strang_rnaas}, this is needed to get second-order convergence.  In our \amrex-Astro \microphysics\ library \citep{microphysics},
we use C++ templating heavily to reduce the computational
expense of the EOS calls by only computing the thermodynamic quantities needed by the integrator.

Detonations can be difficult to model numerically:
the explosive energy released from reactions occurs
on a much shorter timescale than
the sound-crossing time of a zone.  Many
simulations use a simulation timestep based on the reaction
timescale.  For example, \citet{prometheus} advocated for two constraints, one that limits the change in $T$ over a timestep and another that limits the change of each mass fraction, $X_k$ over a timestep.  These can both be much more
restrictive than the hydrodynamics timescale.  Many current works instead use a reaction timestep based on the change in internal energy, $e$, over the timestep. 
 This is an option in \flash.  This energy-based reaction timestep limiter
 is used, for example, in the double detonation simulations of \citet{kushnirkatz:2020,rivas:2022}.  This helps keep the reactions
and hydrodynamics coupled, but can greatly increase the expense of the
simulation.  It is also the case that since the timestep is so small, $T$ does not change much over the simulation timestep, so evolving only the mass fractions may be reasonable.  
Artificially limiting the
rates \citep{kushnir:2013, shen:2018, kushnirkatz:2020, boos:2021} is also sometimes employed to prevent the reactions from from releasing energy faster than the sound crossing timescale of a cell.  Once the detonation ignition region is spatially resolved, this should not be needed.  As we will see in the simulations presented here, if the reactions explicitly know about the advection that is taking place during the burn, neither of these limiting techniques appear to be needed.

In this paper, we focus on nucleosynthesis, exploring how reactions and hydrodynamics are coupled and the role of tolerances, the ignition of the second detonation,  and the
efficiency of the overall method.  We also look at how the results behave with resolution.  Finally, we discuss some of the difficulties encountered during the simulations.

%======================================================================
% Results
%======================================================================
\section{Simulation Setup}\label{Sec:setup}

\subsection{Numerical method}

We use the freely-available \castro\ simulation code for all the
simulations shown here.  
\castro\ solves the compressible Euler equations, Eqs. \ref{eq:euler1}--\ref{eq:euler4}, using an unsplit PPM algorithm for advection \citep{ppmunsplit,millercolella:2002,ppm}, with
the Riemann solver from \citet{colellaglaz:1985} and the general
stellar equation of state of \citet{timmes_swesty:2000}.  \castro\ also advects
internal energy together with total energy, in the dual-energy formalism introduced in \citet{bryan:1995} (see \citealt{wdmergerI} for details on the implementation).  In
all simulations, the CFL constraint is the
only timestep constraint used.  Self-gravity is done using a full
Poisson solve using geometric multigrid, with Dirichlet boundary
conditions on the domain boundary computed via a multipole expansion
with a maximum order of 6.  \castro\ uses adaptive mesh refinement (AMR) via the \amrex\ library
\citep{amrex_joss}, with 
subcycling throughout the AMR hierarchy, so finer grids are advanced at a smaller timestep than the
coarse grids.

In \citet{castro_simple_sdc}, we introduced a method based on the
ideas of spectral deferred corrections \citep{dutt:2000, bourlioux:2003} for coupling reactions and
hydrodynamics.  We termed the method ``simplified-SDC.''  In
the simplified-SDC method, the overall
time-integration is done iteratively, with the hydrodynamics seeing an
explicit reactive source and the reaction update evolving an ODE
system that includes a piecewise-constant-in-time advective source,
$\Adv{\Uc}^{n+1/2}$,
\begin{equation}
\label{eq:simplesdc}
\ddt{\Uc} = \Rb(\Uc) + \Adv{\Uc}^{n+1/2}
\end{equation}
Here, $\Uc$ is the conserved state, $\Uc = (\rho, \rho \Ub, \rho E,
\rho X_k)^\intercal$, $\Rb(\Uc)$ are the reactive sources, and
$\Adv{\Uc}^{n+1/2}$ is an approximation to the advective update over
the timestep, as computed by the corner transport upwind formulation
of the piecewise parabolic method. Crucially, this advective
term includes an explicit reaction source term and, likewise, the reaction update,
Eq.~\ref{eq:simplesdc}, means that the reactions will ``see'' what
advection is doing as the reaction network is integrated.

The reaction system is integrated using a version of the VODE ODE
integrator \citep{vode} ported to C++.  Our modifications to this
integrator are described in \citet{castro_simple_sdc}.  We use an
analytic approximation to the Jacobian (the main approximation is that
the species derivatives of the screening function are not included)
and found that we get the best results when we disable the caching of the
Jacobian in the integrator.

In addition to the time-integration strategy, integrating the reaction
ODE system requires specifying tolerances, usually both an absolute, $\epsilon_\mathrm{abs}$, and relative tolerance, $\epsilon_\mathrm{rel}$, that are combined together into a weight of the form:
\begin{equation}
w_i = \epsilon_{\mathrm{abs},i} + \epsilon_{\mathrm{rel},i} |y_i|
\end{equation}
where the $y_i$ is one of the variables being integrated by the ODE
integrator and $i$ its the index, indicating that different tolerances
can be used for each variable.  Traditionally one set of tolerances is
used for the mass fractions and another for the energy (if it is
integrated).  We will refer to the species tolerances as $\epsilon_\mathrm{abs}(X_k)$ and $\epsilon_\mathrm{ref}(X_k)$ and the energy tolerances as $\epsilon_\mathrm{abs}(e)$ and $\epsilon_\mathrm{rel}(e)$.  Unfortunately, the tolerances used by simulations are
not normally reported in papers, and in many cases, may not be easily
controlled by a code user at runtime. Some examples can be found in publicly
available networks.  For example, the freely available version of the
{\tt aprox13} network uses $\epsilon_\mathrm{rel} = 10^{-5}$ and a
floor on the mass fractions corresponding to $\epsilon_\mathrm{abs} =
10^{-6}$.  We caution that different integrators
may use different error measures or norms so comparing the tolerance values across integrators can be difficult. 
 Since the simplified-SDC update works in terms of the
conserved variables, $\Uc$, we apply a density weighting to the absolute tolerances when comparing to the values used in Strang integration (see \citealt{castro_simple_sdc}).

All simulations use the same reaction network: the {\tt subch\_simple}
as described in \cite{zhi2023}.  This includes 22 nuclei and 94 rates
from ReacLib \citep{reaclib} and is produced with
\pynucastro\ \citep{pynucastro2}.  Importantly, we include
$\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha,p)\isotm{Ne}{21}$
which creates protons to allow for
$\isotm{C}{12}(p,\gamma)\isotm{N}{13}(\alpha,p)\isotm{O}{16}$.  This
sequence can be faster than
$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$, as pointed out by
\citet{shenbildsten} and is important for getting the detonation speed
correct.  To reduce the size of the network, we have
\pynucastro\ combine some $(\alpha,p)(p,\gamma)$ and $(\alpha,\gamma)$
rates into an effective $(\alpha,\gamma)$ rate.  We also approximate
the neutron captures for $\isotm{C}{12}(\isotm{C}{12},\isotm{n}{})\isotm{Mg}{23}(\isotm{n}{},\gamma)\isotm{Mg}{24}$, $\isotm{O}{16}(\isotm{C}{12},\isotm{n}{})\isotm{S}{31}(\isotm{n}{},\gamma)\isotm{S}{32}$, $\isotm{O}{16}(\isotm{O}{16}, \isotm{n}{})\isotm{S}{31}(\isotm{n}{},\gamma)\isotm{S}{32}$, eliminating
the intermediate nucleus in each sequence.  Together these approximations make the 22 nuclei network approximate a 31 nuclei network.  Screening is provided
following the procedure in \citet{wallace:1982}, combining the
screening functions of \citet{graboske:1973,alastuey:1978,itoh:1979}.

Our choice of reaction network size falls in the middle of those reported in the literature.  \citet{townsley:2019,boos:2021} use a 55-isotope network, but do not describe whether energy / temperature is integrated together with the rates (although, based on \citealt{townsley:2016} it likely is not). 
\citet{gronow:2020,gronow:2021} use more than 30 isotopes, and use the integration method from \citet{pakmor:2012}. 
\citet{rivas:2022} use the 13-isotope {\tt aprox13} network, and likewise do not state whether energy / temperature is integrated together with the rates, although \flash\ does not usually include energy/temperature evolution with the reactions.
\citet{mollwoosley:2013} used the {\tt aprox19} network together with a table from a large network to better describe the nucleosynthesis.   While they used \castro,
the temperature evolution during the reactions was simpler then than in the current versions of \castro.  \cite{roy:2022} also use {\tt aprox19}.  
The {\tt aprox13} and {\tt aprox19} networks do not have the reaction sequences discussed in \citet{shenbildsten} and will lead to an underestimate of the He detonation speed, which in turn affects
where the compressiion ignites the C detonation.
None of these works discuss the integration tolerances.  Several works also disable burning in shocks (this includes \citealt{kushnirkatz:2020,gronow:2020,gronow:2021,rivas:2022} and likely others as this can be the default behavior in some codes).  We do not take that approach here in part because the burning will explicitly see what is happening from advection as we integrate the network (see also \citealt{papatheodore:2014}).

If during the advance of a timestep an
error is generated (negative density, ODE integration fails because it cannot meet the tolerances required or takes too many steps, mass
fractions don't sum to 1, CFL constraint violated at the new time),
the step is thrown-out and retried with a smaller timestep.  This is
done on a level-by-level basis in the overall AMR subcycling
hierarchy.  This is discussed further in \citet{castro_simple_sdc}.

\subsection{Initial model}

We considered an extreme version of the double detonation scenario in
\citet{castro_simple_sdc}---a very large perturbation was applied to
drive a detonation directly in a pure carbon
white dwarf.  We showed that the Strang-splitting integration method (including energy evolution) had
difficulty with the integration unless we used tighter tolerances,
while the simplified-SDC method was well-behaved.  Here we look at a
much more realistic setup for a double detonation
and look at convergence with integration method and resolution.  Our new model uses a much smaller
perturbation that drives a He detonation around the C/O white dwarf and a
compression wave that results in the ignition of a C detonation near
the center of the star.

We will run a suite of simulations all using the same initial model,
constructed following the methodology in \citet{subchandra}.  We use
an isothermal core of C/O of $1.1~M_\odot$ with $T = 10^7~\mathrm{K}$
and a thin transition region where the temperature ramps up to
$1.75\times 10^8~\mathrm{K}$ and the composition changes to 99\%
\isot{He}{4} and 1\% \isot{N}{14}.  This He envelope is then
integrated isentropically.  The density of the transition from the
underlying CO white dwarf to the He envelope was selected to yield an
envelope mass of $0.05~M_\odot$.  The code for generating the model is
freely available \citep{initial_models}.
Our choice of white dwarf mass and
envelope mass was inspired by our previous work on modeling the convective runaway in double detonations \citep{jacobs:2016}, where we found larger envelopes and/or core masses are needed to get the localized runaway that could lead to a double detonation.

We place a small temperature perturbation in the He layer using the
same form as in \citet{castro_simple_sdc}:
\begin{equation}
  T = T_0 \left \{ 1 + X(\isotm{He}{4}) f \left [1 + \tanh(2 - \sigma) \right ] \right \}
\end{equation}
where
\begin{equation}
  \sigma = \left [ x^2 + (y - R_0)^2 \right ]^{1/2} / \lambda
\end{equation}
and
\begin{equation}
  R_0 = R_\mathrm{pert} + R_\mathrm{base}
\end{equation}
where, $R_\mathrm{base}$ is the radius at which the helium layer
begins (determined as the radius where $X(\isotm{He}{4})$ first goes beyond $0.5$) and $R_\mathrm{pert}$ is the distance above the base to put the
perturbation.  We choose $R_\mathrm{pert} = 100~\mathrm{km}$.  The
temperature is perturbed above the initial model value, denoted as
$T_0$ here.  The amplitude of the perturbation is $f = 3$ and the
scale of the perturbation is $\lambda = 12.5~\mathrm{km}$.  This is a
very small perturbation, with a peak temperature of about $1.1\times
10^9~\mathrm{K}$, spread over roughly 2 zones (in our references 20 km resolution simulations).  This
is large enough to seed the initial He detonation in the
envelope.

\begin{figure}
\centering
\plotone{subch_Temp_zoom}
\caption{\label{fig:initial_hotspot} Zoom-in on the initial hotspot 100 km above the He-C interface.  The dotted contours show densities
of $10^4$, $10^5$, and $10^6~\gcc$}
\end{figure}

\subsection{Suite of Runs}
